---
title: "Presentation:"
subtitle: "CHAPTER 4. DISTRIBUTIONS OF RANDOM VARIABLES"
author: "Gabriel Campos"
date: "9/27/2020"
output:
  pdf_document: default
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('DATA606')
library(tidyverse)
library(openintro)
```

# Question 4.3 pg.142

## GRE scores, Part I. 

Sophia who took the Graduate Record Examination (GRE) scored **160** on the Verbal Reasoning section and **157** on the Quantitative Reasoning section. The mean score for Verbal Reasoning section for **all** test takers was **151** with a **standard deviation** of **7**, and the mean score for the Quantitative Reasoning was **153** with a **standard deviation** of **7.67**. Suppose that both distributions are nearly normal.


## (a)
Write down the short-hand for these two normal distributions.\newline
*Reference Section 4.1 pg. 133*\newline
\
**Verbal reasoning section: N($\mu$ = 151, $\sigma$ = 7)** \newline
**Quantitative Reasoning section: N($\mu$ = 153, $\sigma$ = 7.67)**\newline

## (b)
### (i)
What is Sophia’s Z-score on the Verbal Reasoning section?\newline 
*Reference 4.1.2 pg. 134* \newline
\
**Z = $\frac{x-\mu}{\sigma}$ Where $x_{verbal}$=160** \newline
```{r}
(160-151)/7
```
### (ii)
On the Quantitative Reasoning section?\newline
**Where $x_{quantitive}$=157** \newline
```{r}
(157-153)/7.67
```

### (iii)
Draw a standard normal distribution curve and mark these two Z-scores.\newline

```{r Verbal, echo=FALSE}
pnorm(1.285714)
normalPlot(bound = c(-Inf,1.285714))
```

--------------------------------------------------------------------------------

\clearpage

### (iv)

```{r Quantitive, echo=FALSE}
pnorm(0.5215124)
normalPlot(bound = c(-Inf,0.5215124))
```

## (c)
What do these Z-scores tell you?\newline
**The Z-scores tell us how far above the mean Sophia scored for each section.**\newline
**For the verbal section Sophia scored 1.285714 standard deviations above the mean.**\newline
**For the quantitative section Sophia scored 0.5215124 standard deviations above the mean.**\newline

## (d)
Relative to others, which section did she do better on?\newline
**Sophia's Z-score for the Verbal Reasoning Section is higher indicating she did better on that section relative to others** \newline

--------------------------------------------------------------------------------

\clearpage

## (e)
Find her percentile scores for the two exams.
*Using normalPlot() or pnorm() results about*\newline

**Verbal Reasoning Section: 90th percentile** \newline
**Quantitative Reasoning Section: 70th percentile** \newline
\

## (f)
### (i)
What percent of the test takers did better than her on the Verbal Reasoning section?\newline
*Using 1-pnorm(Z-score) where one represents 100% of the population minus Sophia's percentile ranking*\newline
*Or subtracting percentages provided e.g. 100% - 90% = 10%* \newline
**10% of the test takers did better than Sophia on Verbal Reasoning Section**\newline
```{r}
1-pnorm(1.285714)
```
### (ii)
On the Quantitative Reasoning section?\newline
*Using 1-pnorm(Z-score) where one represents 100% of the population minus Sophia's percentile ranking*\newline
*Or subtracting percentages provided e.g. 100% - 70% = 30%* \newline
**30% of the test takers did better than Sophia on Verbal Reasoning Section**\newline
```{r}
1-pnorm(0.5215124)
```

## (g)
Explain why simply comparing raw scores from the two sections could lead to an incorrect conclusion as to which section a student did better on.\newline
**The weight of Sophia's raw score is on a different scaling system, so comparing her performance against the population makes sense in concluding how well she did.** \newline

## (h)
If the distributions of the scores on these exams are not nearly normal, would your answers to parts (b)-(f) change? Explain your reasoning.\newline
**Z-scores will still be calculated with the same formula, but without a nearly normal distribution, we would not be able to use the normal probability table. Therefore we could answer (b)-(c) but not (d)-(f).**

