---
title: DATA 606 Data Project Proposal
author: Gabriel Campos
output:
  pdf_document: default
---

### Data Preparation

```{r setup, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
# load data
library(tidyverse)
library(openintro)
library(infer)
library(dplyr)
library(VennDiagram)
library(scales)
library(data.table)
library (readr)
project_data<- read_csv("D:/Bulk Data/database.csv")

```

```{r regional-subsets, echo=TRUE, results= FALSE, include=TRUE}
west <- project_data %>%
  filter(State == "Arizona"| State == "California"| State =="Colorado"
  	   | State =="Idaho"| State =="Montana"| State =="Nevada"
  	   | State == "New Mexico"| State =="Oregon" | State == "Utah"
  	   | State =="Washington"| State =="Wyoming")
midwest <- project_data %>%
  filter(State == "Illinois"| State == "Indiana"| State =="Iowa"
  	   | State =="Kansas"| State =="Michigan"| State =="Minnesota"
  	   | State == "Missouri"| State =="Nebraska"| State == "North Dakota"
  	   | State =="Ohio"| State =="South Dakota"| State =="Wisconsin")
northeast <- project_data %>%
  filter(State == "Connecticut"| State == "Delaware"| State =="Maine"
  	   | State =="Massachusetts"| State =="New Hampshire"
  	   | State =="New Jersey"| State == "New York"| State =="Pennsylvania"
  	   | State == "Rhode Island"| State =="Vermont")
south <- project_data %>%
  filter(State == "Alabama"| State == "Arkansas"| State =="Florida"
  	   | State =="Georgia"| State =="Kentucky"| State =="Louisiana"
  	   | State == "Mississippi"| State =="North Carolina"| State == "Oklahoma"
  	   | State =="South Carolina"| State =="South Carolina"| State =="Tennessee"
  	   | State =="Texas"| State =="Virgina"| State =="West Virgina")
pacific<- project_data %>%
  filter(State == "Alaska"| State == "Hawaii")
```

![US regions based on [census.gov](https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf) ](regions.PNG){height=25%}


### Research question 

**Question: Between 1980-2014, was the South the most dangerous US region for women to live in?**

$H_0$ = $Region_{south.30yrs.OR.less}\le Region_{(west...midwest...northeast...pacific).30yrs.OR.less}$\newline
$H_A$ = $Region_{south.30yrs.OR.less}> Region_{(west...midwest...northeast...pacific).30yrs.OR.less}$\newline

### Cases 

**What are the cases, and how many are there?**

**Categorical Data:** Record ID, Agency Code, Agency Name, Agency Type, City, State,Incident,\newline
Crime Type,Crime Solved,Victim Sex,Victim Race,Victim Ethnicity,Perpetrator Sex,\newline
Perpetrator Race,Perpetrator,Ethnicity,Relationship,Weapon,Record Source, \newline
Region (**Derived by creating sub datasets NOTE CHUNK $r$ $regional-subsets$**) \newline
\
**Numerical:** Year,Month,Victim Age,Perpetrator Age,Victim Count,Perpetrator Count \newline
\
**Outliers:** 998 \newline
\
There are 66,301 total cases in our data set, representing murders committed against Female's under the age of 30 throughout the United States from 1980-2014.\newline

```{r, echo = TRUE, results='hide'}
project_data %>% filter(`Victim Sex` == "Female" & `Victim Age`< 31) %>% glimpse()
```


```{r, echo=FALSE, include=FALSE}
glimpse(project_data)
paste('West coast cases ',NROW(na.omit(west)))
paste('Midwest cases ',NROW(na.omit(midwest)))
paste('Northeast cases ',NROW(na.omit(northeast)))
paste('South cases ' , NROW(na.omit(south)))
paste('Pacific cases', NROW(na.omit(pacific)))
```




### Data collection 

**Describe the method of data collection.**\newline

*Data source: [kaggle.com](https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv)*\newline
	
	Content
	The Murder Accountability Project is the most complete database of homicides in 
	the United States currently available. This dataset includes murders from the FBI's
	Supplementary Homicide Report from 1976 to the present and Freedom of Information
	Act data on more than 22,000 homicides that were not reported to the Justice Department.
	This dataset includes the age, race, sex, ethnicity of victims and perpetrators, in
	addition to the relationship between the victim and perpetrator and weapon used.

	Acknowledgements
	The data was compiled and made available by the Murder Accountability Project, founded by 
	Thomas Hargrove.
	
*Homicide Reports, 1980-2014
Project
https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv*


### Type of study 

**What type of study is this (observational/experiment)?**\newline
\
Observational


### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

*Homicide Reports, 1980-2014
Project
https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv*


### Dependent Variable

```{r, echo=FALSE}
names(project_data)
```


**What is the response variable? Is it quantitative or qualitative?**

Murder count, which is counted with Record ID and is quantitative.

### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**\newline
Victim Age (Quantitative) and Victim Sex (Qualitative)

------------------------------------------------------------------------------------------------------------
\clearpage

### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r hist_west_coast, echo=FALSE, fig.width=4, fig.height=3, include=FALSE}
# option 2 for plot

plot( as.factor(west$Year),main ='West Coast Murders 1980-2014' , las=2) 
plot( as.factor(midwest$Year),main ='Midwest Coast Murders 1980-2014' , las=2)
plot( as.factor(northeast$Year),main ='Northeast Coast Murders 1980-2014' , las=2)
plot( as.factor(south$Year),main ='South Coast Murders 1980-2014' , las=2)
plot( as.factor(pacific$Year),main ='Pacific Coast Murders 1980-2014' , las=2)
```

```{r, results='hide'}
glimpse(project_data)
```

**There were 638,454 total murders between 1980-2014**
\

```{r, echo=FALSE}
plot( as.factor(project_data$Year),main ='US Murders 1980-2014' , las=2)

```



\clearpage


**143,345 of the murders were specifically against Women**

```{r, echo = TRUE, results='hide'}
project_data %>% filter(`Victim Sex` == "Female") %>% glimpse()
```


```{r, echo=FALSE, fig.width=4, fig.height=3}
ggplot(project_data)+geom_bar(aes(x=`Victim Sex`))

```



**A break down of murders by age are as follows $Note:$ $age$ $998$ $was$ $assigned$ $to$ $victims$ $who's$ $age$ $could$ $not$ $be$ $determined$**

```{r, echo=FALSE}
count<-project_data$`Victim Age`
table(count)
```
\clearpage

```{r}
labs <- c("1-10","11-20","21-30","31-40","41-50","51-60","61-70","71-80","81-90","91-100")
#labs
project_data$`Victim AgeGroup`<- cut(project_data$`Victim Age`,
									 breaks=c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100),
									 right = FALSE)
ggplot(project_data)+geom_bar(aes(x=`Victim AgeGroup`))
```


