---
title: Relationship between US Regions and Homocide Rates of Young Women between 1980-2014
subtitle: DATA 606 Data Project Submission
author: Gabriel Campos
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
paste0(" Attaching packages: DATA606,infer,dplyr,VennDiagram,scales,data.table,readr ")
```

```{r echo=FALSE, warning=FALSE, message=FALSE,include=FALSE}
library(DATA606)
library(infer)
library(dplyr)
library(VennDiagram)
#library(scales)
library(data.table)
library (readr)
```


## Part 1 - Introduction

	A primary concern regarding public safety that, impacts all of society is murder.
	The threat of murder negatively impacts the widely accepted fundamental human need
	for safety.	Murder by region can impact the population psychologically, economically 
	and can stifle a community's continued progress for a better standard of living.
	Gaining insight on those who are the most vulnerable and impacted by murder is 
	necessary to forge initiatives in assisting those victimized by its threat. Like many
	approaches in troubleshooting or diagnosing an issue, understanding where to begin is
	a key first step. In order to focus this analysis, we will focus specifically on murder
	rates for 'women under 30' against the entire data set.
	
### Overview

	The goals for this project are to:

* Think about the independent and dependent variables in correlation to murder rates including region, gender and age
* Compare murder rates by region, race, gender and age based on the data set
* Draw conclusions of the most at risk region for our specific population in the United States based on our findings

\

***
\clearpage
	
## Part 2 - Data

### Collection

	The Murder Accountability Project is the most complete database of homicides in 
	the United States currently available. This dataset includes murders from the FBI's
	Supplementary Homicide Report from 1976 to the present and Freedom of Information
	Act data on more than 22,000 homicides that were not reported to the Justice Department.
	This dataset includes the age, race, sex, ethnicity of victims and perpetrators, in
	addition to the relationship between the victim and perpetrator and weapon used. A
	victims age is rounded down by year (e.g. toddlers 11 months old or younger qualify as 0 years old).
	If a victim's age cannot be determined, they will be categorized as 998 years old
	respectively. 
*Data source: [kaggle.com](https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv)*\newline

### Load Data

```{r vecotr-region, echo=FALSE, message=FALSE, warning=FALSE}
project_data<- read_csv("D:/Bulk Data/database.csv")
project_data$Region<-
	ifelse(project_data$State=="Arizona" |
		   	project_data$State=="California"| project_data$State=="Colorado"| project_data$State=="Idaho"|
		   	project_data$State=="Montana"| project_data$State=="Nevada"| project_data$State=="New Mexico"|
		   	project_data$State=="Oregon"| project_data$State=="Utah"| project_data$State=="Washington"|
		   	project_data$State=="Wyoming","West",
		   ifelse(project_data$State=="Illinois"|
		   	   	project_data$State == "Indiana"| project_data$State == "Iowa"| project_data$State == "Kansas"|
		   	   	project_data$State == "Michigan"| project_data$State == "Minnesota"| project_data$State ==
		   	   	"Missouri"| project_data$State == "Nebraska"| project_data$State == "North Dakota"|
		   	   	project_data$State == "Ohio"| project_data$State == "South Dakota"| project_data$State ==
		   	   	"Wisconsin","Midwest",
		   	   ifelse(project_data$State == "Connecticut"| project_data$State == "Delaware"| project_data$State
		   	   	   =="Maine"| project_data$State =="Massachusetts"| project_data$State =="New Hampshire"
		   	   	   |project_data$State =="New Jersey"| project_data$State == "New York"| project_data$State
		   	   	   =="Pennsylvania"| project_data$State == "Rhode Island"| project_data$State=="Vermont","Northeast",
		   	   	   ifelse(project_data$State == "Alaska"| project_data$State == "Hawaii","Pacific","South")
		   	   	   )
		   	   )
		   )

project_data$`Victim AgeGroup`<- cut(project_data$`Victim Age`,
									 breaks=c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100),
									 right = FALSE)

```



```{r, echo=FALSE, warning=FALSE, message=FALSE}
west <- project_data %>%
  filter(State == "Arizona"| State == "California"| State =="Colorado"
  	   | State =="Idaho"| State =="Montana"| State =="Nevada"
  	   | State == "New Mexico"| State =="Oregon" | State == "Utah"
  	   | State =="Washington"| State =="Wyoming")
midwest <- project_data %>%
  filter(State == "Illinois"| State == "Indiana"| State =="Iowa"
  	   | State =="Kansas"| State =="Michigan"| State =="Minnesota"
  	   | State == "Missouri"| State =="Nebraska"| State == "North Dakota"
  	   | State =="Ohio"| State =="South Dakota"| State =="Wisconsin")
northeast <- project_data %>%
  filter(State == "Connecticut"| State == "Delaware"| State =="Maine"
  	   | State =="Massachusetts"| State =="New Hampshire"
  	   | State =="New Jersey"| State == "New York"| State =="Pennsylvania"
  	   | State == "Rhode Island"| State =="Vermont")
south <- project_data %>%
  filter(State == "Alabama"| State == "Arkansas"| State =="Florida"
  	   | State =="Georgia"| State =="Kentucky"| State =="Louisiana"
  	   | State == "Mississippi"| State =="North Carolina"| State == "Oklahoma"
  	   | State =="South Carolina"| State =="South Carolina"| State =="Tennessee"
  	   | State =="Texas"| State =="Virgina"| State =="West Virgina")
pacific<- project_data %>%
  filter(State == "Alaska"| State == "Hawaii")

paste0("Loading Data: database.csv ")
paste0("Subset Data: Region <- project_data %>% filter() ")
paste0("Vector Region: project_data$Region<-ifelse(State,Region1,ifelse(State,Region2,")
paste0("Vector AgeGroup: project_data$`Victim AgeGroup`<- cut(project_data$`Victim Age`,... ")
```

![US regions based on [census.gov](https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf) ](regions.PNG){height=40%}



\clearpage

### Cases

**Categorical Data:** Record ID, Agency Code, Agency Name, Agency Type, City, State,Incident,\newline
Crime Type,Crime Solved,Victim Sex,Victim Race,Victim Ethnicity,Perpetrator Sex,\newline
Perpetrator Race,Perpetrator Ethnicity,Relationship,Weapon,Record Source, \newline
Region\newline
\
**Numerical:** Year,Month,Victim Age,Perpetrator Age,Victim Count,Perpetrator Count \newline
\
**Outliers:** 974 \newline

```{r, echo = FALSE, results='hide'}
project_data %>% filter(`Victim Sex` == "Female" & `Victim Age`< 31) %>% glimpse()
glimpse(project_data)
```

There are **638,454** total cases in our data set, with **66,301** representing **murders** committed against **Female's under** the age of **30** throughout the United States from 1980-2014

**A break down of murders by age are as follows**\newline
$Note:Age\ 998\ was\ assigned\ to\ victims\ who's\ age\ could\ not\ be\ determined.$\newline
$To\ avoid\ the\ impact\ it\ will\ have\ on\ our\ results\ I\ will\ remove\ it\ from\ our\ existing\ dataframes$

```{r, echo=FALSE}
count<-project_data$`Victim Age`
table(count)
project_data<-project_data[!(project_data$`Victim Age`==998) & project_data$`Victim Sex`!='Unknown',]
west<-west[!(west$`Victim Age`==998) & west$`Victim Sex`!='Unknown',]
midwest<-midwest[!(midwest$`Victim Age`==998) & midwest$`Victim Sex`!='Unknown',]
northeast<-northeast[!(northeast$`Victim Age`==998) & northeast$`Victim Sex`!='Unknown',]
south<-south[!(south$`Victim Age`==998) & south$`Victim Sex`!='Unknown',]
pacific<-pacific[!(pacific$`Victim Age`==998) & pacific$`Victim Sex`!='Unknown',]
paste0("Non determined variables removed with the following command:")
paste0("data.frame<-data.frame[!(dataframe$`Victim Age`==998),]")

```
\clearpage

	The charts below show that of the murders reported, the majority of these are against men.
	Additionally, regardless of gender, nationally the highest murder victims age falls with 
	Victims age 20-30.

```{r, echo=FALSE,fig.height=3}
ggplot(project_data)+geom_bar(aes(x=`Victim Sex`,fill = `Victim Sex`))+coord_flip()+scale_fill_hue(l=40)
ggplot(project_data,aes(x = `Victim AgeGroup`,fill = `Victim AgeGroup`))+geom_bar()+scale_fill_hue(l=40)
```

***

\clearpage

## Part 3 - Exploratory data analysis

### National Age Average

```{r, echo=FALSE}
paste0("Summary of age for entire dataset")
summary(project_data$`Victim Age`)
sd(project_data$`Victim Age`)
```
\

* **Murders 1980-2014: N($\mu$ = 33.26, $\sigma$ $\approx$ 17.79)**
* I would like to see the probability of a murder taking place on `Victim Sex` $\le$ 30
* In order to do so I can calculate **Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.183249$

```{r, echo=FALSE, results='hide',fig.height=3}
(30-33.26)/17.79
pnorm( -0.183249)
normalPlot(bound = c(-Inf, -0.183249))
```

What's interesting here, $\approx$  $42.7\%$ of all the reported murders in our data set, of victims\newline
whose age we can determine, occurs with the first 30 years of there life. Next I will\newline
compare the average age of murdered women in our dataset, by region.


```{r create new var full, echo=FALSE}
#WEST
project_data <- project_data %>% 
  mutate(femW = ifelse(project_data$Region=="West" & project_data$`Victim Sex`=="Female", "yes", "no"))


#MIDWEST
project_data <- project_data %>% 
  mutate(femMid = ifelse(project_data$Region=="Midwest" & project_data$`Victim Sex`=="Female", "yes", "no"))

#NORTHEAST
project_data <- project_data %>% 
  mutate( femNE= ifelse(project_data$Region=="Northeast" & project_data$`Victim Sex`=="Female", "yes", "no"))

#SOUTH
project_data <- project_data %>% 
  mutate( femS= ifelse(project_data$Region=="South" & project_data$`Victim Sex`=="Female", "yes", "no"))

#PACIFIC
project_data <- project_data %>% 
  mutate( femP= ifelse(project_data$Region=="Pacific" & project_data$`Victim Sex`=="Female", "yes", "no"))
```

\clearpage

### Regional Age Averages

```{r, echo=FALSE}
paste0("In order to compare variabilities I created vectors in our master data set using")
paste0("project_data <- project_data %>% mutate(femW = ifelse(project_data$Region....")
paste0("allowing for comparison of one regions variability with the remaining 3 combined")
```

	The boxplots below show that with respect to age, our data is right skewed, indicating 
	murders occurred at a higher frequency within the first three decades of our victims lives.
	This is supported with consistent medians, relatively close means and similar number of 
	outliers shown. The only standouts would be the Pacific, which has slightly more scattered 
	outliers, and the South which has the lowest percentage of murder victims below 30 ( 40% ).
	Based on the data, I suspect age brackets hardly impact murders among women below 30.
	

**Western Murders 1980-2014: N($\mu$ = 32.64, $\sigma$ = 17.42)**\newline
**Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.1515663$ making $\approx$ $44\%$ of murder victims years$\le$ **30**

```{r boxplot west, out.width="50%",echo=FALSE,results='hide'}
summary(west$`Victim Age`)
sd(west$`Victim Age`)
(30-32.64)/17.41812
pnorm(-0.1515663)
normalPlot(bound = c(-Inf,-0.1515663))
ggplot(project_data, aes(y=`Victim Age`, x=femW))+geom_boxplot()+labs(title="Western Female gender vs. Remaining Regions",x="West & Female")
```
\
**Midwestern Murders 1980-2014: N($\mu$ = 32.6, $\sigma$ $\approx$ 17.34)**\newline
**Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.1499721$ making $\approx$ $44\%$ of murder victims years$\le$ **30**

```{r boxplot midwest, out.width="50%",echo=FALSE,results='hide'}
summary(midwest$`Victim Age`)
sd(midwest$`Victim Age`)
(30-32.6)/17.33656
pnorm(-0.1499721)
normalPlot(bound = c(-Inf,-0.1499721))
ggplot(project_data, aes(y=`Victim Age`, x=femMid))+geom_boxplot()+labs(title="MidWestern Female gender vs. Remaining Regions",x="MidWest & Female")
```

\clearpage

**Northeastern Murders 1980-2014: N($\mu$ = 32.99, $\sigma$ $\approx$ 17.56)**\newline
**Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.1702733$ making $\approx$ $43\%$ of murder victims years$\le$ **30**

```{r boxplot northeast, out.width="50%",echo=FALSE,results='hide'}
summary(northeast$`Victim Age`)
sd(northeast$`Victim Age`)
(30-32.99)/17.56
pnorm(-0.1702733)
normalPlot(bound = c(-Inf,-0.1702733))
ggplot(project_data, aes(y=`Victim Age`, x=femNE))+geom_boxplot()+labs(title="NorthEastern Female gender vs. Remaining Regions",x="NorthEast & Female")
```

**Southern Murders 1980-2014: N($\mu$ = 35, $\sigma$ $\approx$ 18.32)**\newline
**Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.2728698$ making $\approx$ $40\%$ of murder victims years$\le$ **30**

```{r boxplot south, out.width="50%",echo=FALSE,results='hide'}
summary(south$`Victim Age`)
sd(south$`Victim Age`)
(30-35)/ 18.32376
pnorm(-0.2728698)
normalPlot(bound = c(-Inf,-0.2728698))
ggplot(project_data, aes(y=`Victim Age`, x=femS))+geom_boxplot()+labs(title="Southern Female gender vs. Remaining Regions",x="Southern & Female")
```

**Pacific Murders 1980-2014: N($\mu$ = 33.81, $\sigma$ $\approx$ 17.12)**\newline
**Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.1168525$ making $\approx$ $45\%$ of murder victims years$\le$ **30**

```{r boxplot pacific, out.width="50%",echo=FALSE,results='hide'}
summary(pacific$`Victim Age`)
sd(pacific$`Victim Age`)
(30-32)/17.11559
pnorm(-0.1168525)
normalPlot(bound = c(-Inf,-0.1168525))
ggplot(project_data, aes(y=`Victim Age`, x=femP))+geom_boxplot()+labs(title="Pacific Female gender vs. Remaining Regions",x="Pacific & Female")

```

***

\clearpage

### Average Female Murder Count by Region

	Next I created a dataframe, categorizing the annual murder count of women by region
	for the years between 1980-2014. What I found appeared alarming at first. I seems the
	South has an overwhelming higher total female murder count:
\
\
\
	
\begin{tabular}{l  r  r  r  r  r  r}
                & \multicolumn{5}{c}{\textit{Female Murder Count by Region}} \\
\cline{2-6}
                & West  & Midwest  & Northeast & South & Pacific & Total \\
\hline
Mean            & 890.2     & 768     & 623.65       & 1,786.25   & 25.37   & 818.6971 \\
SD              & 160.84    & 163.9   & 149.09       & 241.81     & 7.23    & 591.7687 \\
n               & 31,157    & 26,880  & 21,82        & 62,519     & 888     & 143,272 \\
\hline
\end{tabular}
\newline
\newline
**Things to note regarding this data:**

	(1) visually one can suspect the general size of the South region might attribute to
	its numbers
	(2) The South has had the steepest increase in $Murders_{women\ under\ 30 }$ in the 
	early 90's
	(3) The South has had the steepest decline in $Murders_{women\ under\ 30 }$ in the early
	90's to 2014
	(4) $Murders_{women\ under\ 30 }$ in the Pacific has been the most constant among the batch
\



```{r means_female_only, echo=FALSE, results='hide'}
female_only<-data.frame(Year = project_data$Year,Region = project_data$Region, Sex = project_data$`Victim Sex`)
female_only<-female_only[female_only$Sex=='Female',]
female_only<-female_only%>%count(Sex,Region,Year)
female_only%>%filter(Region=="West")%>%summarize(mean(n))
female_only%>%filter(Region=="Midwest")%>%summarize(mean(n))
female_only%>%filter(Region=="Northeast")%>%summarize(mean(n))
female_only%>%filter(Region=="South")%>%summarize(mean(n))
female_only%>%filter(Region=="Pacific")%>%summarize(mean(n))
```

```{r sd_female_only, echo=FALSE, results='hide'}
female_only%>%filter(Region=="West")%>%summarize(sd(n))
female_only%>%filter(Region=="Midwest")%>%summarize(sd(n))
female_only%>%filter(Region=="Northeast")%>%summarize(sd(n))
female_only%>%filter(Region=="South")%>%summarize(sd(n))
female_only%>%filter(Region=="Pacific")%>%summarize(sd(n))
```

```{r sum_female_only, echo=FALSE, results='hide'}
female_only%>%filter(Region=="West")%>%summarize(sum(n))
female_only%>%filter(Region=="Midwest")%>%summarize(sum(n))
female_only%>%filter(Region=="Northeast")%>%summarize(sum(n))
female_only%>%filter(Region=="South")%>%summarize(sum(n))
female_only%>%filter(Region=="Pacific")%>%summarize(sum(n))
```

```{r, echo=FALSE, results='hide'}
mean(female_only$n)
sd(female_only$n)
sum(female_only$n)
project_data %>% filter(`Victim Sex` == "Female" & `Victim Age`< 31) %>% count()
```

```{r, echo=FALSE, out.height="50%"}

ggplot(data = female_only, aes(x = Year, y = n, color=Region, group = Region, )) +
  geom_line() +
	geom_point()+
	ggtitle("Annual Female Murder Count by Region")+
	ylab("Muder Count")
```

```{r echo=FALSE, warning=FALSE}
paste0("Data Frame: female_only<-data.frame(Year = project_data$Year,R...")
paste0( " [female_only$Sex!='Male',] ...count(Sex,Region,Year)  ")
```

## Part 4 - Inference


## Part 5 - Conclusion


## References:

### Homicide Reports, 1980-2014

*Project
https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv*

## Appendix (optional):

Remove this section if you don't have an appendix
