---
title: Relationship between US Regions and Homocide Rates of Young Women between 1980-2014
subtitle: DATA 606 Data Project Submission
author: Gabriel Campos
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
paste0(" Attaching packages: DATA606,infer,dplyr,VennDiagram,scales,data.table,readr ")
```

```{r echo=FALSE, warning=FALSE, message=FALSE,include=FALSE}
library(DATA606)
library(infer)
library(dplyr)
library(VennDiagram)
#library(scales)
library(data.table)
library (readr)
```


## Part 1 - Introduction

	A primary concern regarding public safety that, impacts all of society is murder.
	The threat of murder negatively impacts the widely accepted fundamental human need
	for safety.	Murder by region can impact the population psychologically, economically 
	and can stifle a community's continued progress for a better standard of living.
	Gaining insight on those who are the most vulnerable and impacted by murder is 
	necessary to forge initiatives in assisting those victimized by its threat. Like many
	approaches in troubleshooting or diagnosing an issue, understanding where to begin is
	a key first step. In order to focus this analysis, we will look further into what 
	attributes impact the murder counts in the United States based on the data set.
	
### Overview

	The goals for this project are to:

* Think about the independent and dependent variables in correlation to murder counts including region, gender and age.
* Compare murder rates by region, gender and age based on the data set
* Draw conclusions of the factors impacting the counts and question if these factors are localized based on our findings.

\

***
\clearpage
	
## Part 2 - Data

### Collection

	The Murder Accountability Project is the most complete database of homicides in 
	the United States currently available. This dataset includes murders from the FBI's
	Supplementary Homicide Report from 1976 to the present and Freedom of Information
	Act data on more than 22,000 homicides that were not reported to the Justice Department.
	This dataset includes the age, race, sex, ethnicity of victims and perpetrators, in
	addition to the relationship between the victim and perpetrator and weapon used. A
	victims age is rounded down by year (e.g. toddlers 11 months old or younger qualify as 0 years old).
	If a victim's age cannot be determined, they will be categorized as 998 years old
	respectively. 
*Data source: [kaggle.com](https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv)*\newline

### Load Data

```{r vecotr-region, echo=FALSE, message=FALSE, warning=FALSE}
project_data<- read_csv("D:/Bulk Data/database.csv")
project_data$Region<-
	ifelse(project_data$State=="Arizona" |
		   	project_data$State=="California"| project_data$State=="Colorado"| project_data$State=="Idaho"|
		   	project_data$State=="Montana"| project_data$State=="Nevada"| project_data$State=="New Mexico"|
		   	project_data$State=="Oregon"| project_data$State=="Utah"| project_data$State=="Washington"|
		   	project_data$State=="Wyoming","West",
		   ifelse(project_data$State=="Illinois"|
		   	   	project_data$State == "Indiana"| project_data$State == "Iowa"| project_data$State == "Kansas"|
		   	   	project_data$State == "Michigan"| project_data$State == "Minnesota"|
		   	   	project_data$State == "Missouri"| project_data$State == "Nebraska"|
		   	   	project_data$State == "North Dakota"| project_data$State == "Ohio"|
		   	   	project_data$State == "South Dakota"| project_data$State == "Wisconsin","Midwest",
		   	   ifelse(project_data$State == "Connecticut"|
		   	   	    project_data$State =="Maine"|project_data$State =="Massachusetts"|
		   	   	   	project_data$State =="New Hampshire"| project_data$State =="New Jersey"|
		   	   	   	project_data$State == "New York"| project_data$State =="Pennsylvania"|
		   	   	   	project_data$State == "Rhodes Island"| project_data$State=="Vermont","Northeast",
		   	   	   ifelse(project_data$State == "Alabama"| project_data$State == "Arkansas"|
		   	   	   	   	project_data$State == "Delaware"| project_data$State =="District of Columbia"|
		   	   	   	   	project_data$State == "Florida" | project_data$State == "Georgia"|
		   	   	   	   	project_data$State == "Kentucky"| project_data$State ==	"Louisiana"|
		   	   	   	   	project_data$State == "Maryland"| project_data$State == "Mississippi"|
		   	   	   	   	project_data$State == "North Carolina"| project_data$State == "Oklahoma"|
		   	   	   	   	project_data$State == "South Carolina"| project_data$State == "Tennessee"|
		   	   	   	   	project_data$State == "Texas"| project_data$State == "Virginia"|
		   	   	   	   	project_data$State == "West Virginia","South",
		   	   	   	   ifelse(project_data$State =="Alaska"| project_data$State == "Hawaii", "Pacific",
		   	   	   	   	   "Other"))
		   	   	   )
		   	   )
		   )

project_data$`Victim AgeGroup`<- cut(project_data$`Victim Age`,
									 breaks=c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100),
									 right = FALSE)
count<-project_data$`Victim Age`
project_data<-project_data[!(project_data$`Victim Age`==998) & project_data$`Victim Sex`!='Unknown',]
```



```{r, echo=FALSE, warning=FALSE, message=FALSE}
#west <- project_data %>%
 # filter(State == "Arizona"| State == "California"| State =="Colorado"
  #	   | State =="Idaho"| State =="Montana"| State =="Nevada"
  #	   | State == "New Mexico"| State =="Oregon" | State == "Utah"
  #	   | State =="Washington"| State =="Wyoming")
west<-project_data[project_data$Region=='West',]
paste0("Loading Data: database.csv ")
paste0("Subset Data: Region <- project_data %>% filter() ")
paste0("Vector Region: project_data$Region<-ifelse(State,Region1,ifelse(State,Region2,")
paste0("Vector AgeGroup: project_data$`Victim AgeGroup`<- cut(project_data$`Victim Age`,... ")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#midwest <- project_data %>%
#  filter(State == "Illinois"| State == "Indiana"| State =="Iowa"
#  	   | State =="Kansas"| State =="Michigan"| State =="Minnesota"
#  	   | State == "Missouri"| State =="Nebraska"| State == "North Dakota"
#  	   | State =="Ohio"| State =="South Dakota"| State =="Wisconsin")
midwest<-project_data[project_data$Region=='Midwest',]
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#northeast <- project_data %>%
#  filter(State == "Connecticut"| State == "Delaware"| State =="Maine"
#  	   | State =="Massachusetts"| State =="New Hampshire"
#  	   | State =="New Jersey"| State == "New York"| State =="Pennsylvania"
#  	   | State == "Rhodes Island"| State =="Vermont")
northeast<-project_data[project_data$Region=='Northeast',]
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#south <- project_data %>%
#  filter(State == "Alabama"| State == "Arkansas"| State == "Delaware"
#  	   | State =="District of Columbia" | State =="Florida"
#	   | State =="Georgia"| State =="Kentucky"
#  	   | State =="Louisiana"| State == "Maryland"|State == "Mississippi"
#  	   | State =="North Carolina"| State == "Oklahoma"
#  	   | State =="South Carolina"| State =="Tennessee"
#  	   | State =="Texas"| State =="Virgina"| State =="West Virgina")
south<-project_data[project_data$Region=='South',]
#south %>% filter(`Victim Sex` == "Female") %>% glimpse()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#pacific<- project_data %>%
#  filter(State == "Alaska"| State == "Hawaii")
pacific<-project_data[project_data$Region=='Pacific',]
```









![US regions based on [census.gov](https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf) ](regions.PNG){height=40%}



\clearpage

### Cases

**Categorical Data:** Record ID, Agency Code, Agency Name, Agency Type, City, State,Incident,\newline
Crime Type,Crime Solved,Victim Sex,Victim Race,Victim Ethnicity,Perpetrator Sex,\newline
Perpetrator Race,Perpetrator Ethnicity,Relationship,Weapon,Record Source, \newline
Region\newline
\
**Numerical:** Year,Month,Victim Age,Perpetrator Age,Victim Count,Perpetrator Count \newline
\
**Outliers:** 974 \newline

```{r, echo = FALSE, results='hide'}
project_data %>% filter(`Victim Sex` == "Female" & `Victim Age`< 31) %>% glimpse()
glimpse(project_data)
```

There are **638,454** total cases in our data set, with **66,301** representing **murders** committed against **Female's under** the age of **30** throughout the United States from 1980-2014

**A break down of murders by age are as follows**\newline
$Note:Age\ 998\ was\ assigned\ to\ victims\ who's\ age\ could\ not\ be\ determined.$\newline
$To\ avoid\ the\ impact\ it\ will\ have\ on\ our\ results\ I\ will\ remove\ it\ from\ our\ existing\ dataframes$

```{r, echo=FALSE}

table(count)

west<-west[!(west$`Victim Age`==998) & west$`Victim Sex`!='Unknown',]
midwest<-midwest[!(midwest$`Victim Age`==998) & midwest$`Victim Sex`!='Unknown',]
northeast<-northeast[!(northeast$`Victim Age`==998) & northeast$`Victim Sex`!='Unknown',]
south<-south[!(south$`Victim Age`==998) & south$`Victim Sex`!='Unknown',]
pacific<-pacific[!(pacific$`Victim Age`==998) & pacific$`Victim Sex`!='Unknown',]
paste0("Non determined variables removed with the following command:")
paste0("data.frame<-data.frame[!(dataframe$`Victim Age`==998),]")

```
\clearpage

	The charts below show that of the murders reported, the majority of these are against men.
	Additionally, regardless of gender, nationally the highest murder victims age falls with 
	Victims age 20-30.

```{r, echo=FALSE,fig.height=3}
ggplot(project_data)+geom_bar(aes(x=`Victim Sex`,fill = `Victim Sex`))+coord_flip()+scale_fill_hue(l=40)
ggplot(project_data,aes(x = `Victim AgeGroup`,fill = `Victim AgeGroup`))+geom_bar()+scale_fill_hue(l=40)
```

***

\clearpage

## Part 3 - Exploratory data analysis

### National Age Average

```{r, echo=FALSE}
paste0("Summary of age for entire dataset")
summary(project_data$`Victim Age`)
sd(project_data$`Victim Age`)
```
\

* **Murders 1980-2014: N($\mu$ = 33.56, $\sigma$ $\approx$ 17.79)**
* I would like to see the probability of a murder taking place on `Victim Age` $\le$ 30
* In order to do so I can calculate **Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.2001124$

```{r, echo=FALSE, results='hide',fig.height=3}
(30-33.56)/17.79
pnorm( -0.2001124)
normalPlot(bound = c(-Inf, -0.2001124))
```

What's interesting here, $\approx$  $42.1\%$ of all the reported murders in our data set, of victims\newline
whose age we can determine, occurs with the first 30 years of there life. Next I will\newline
compare the average age of murdered women in our dataset, by region.


```{r create new var full, echo=FALSE}
#WEST
project_data <- project_data %>% 
  mutate(femW = ifelse(project_data$Region=="West" & project_data$`Victim Sex`=="Female", "yes", "no"))


#MIDWEST
project_data <- project_data %>% 
  mutate(femMid = ifelse(project_data$Region=="Midwest" & project_data$`Victim Sex`=="Female", "yes", "no"))

#NORTHEAST
project_data <- project_data %>% 
  mutate( femNE= ifelse(project_data$Region=="Northeast" & project_data$`Victim Sex`=="Female", "yes", "no"))

#SOUTH
project_data <- project_data %>% 
  mutate( femS= ifelse(project_data$Region=="South" & project_data$`Victim Sex`=="Female", "yes", "no"))

#PACIFIC
project_data <- project_data %>% 
  mutate( femP= ifelse(project_data$Region=="Pacific" & project_data$`Victim Sex`=="Female", "yes", "no"))
```

\clearpage

### Regional Age Averages

```{r, echo=FALSE}
paste0("In order to compare variabilities I created vectors in our master data set using")
paste0("project_data <- project_data %>% mutate(femW = ifelse(project_data$Region....")
paste0("allowing for comparison of one regions variability with the remaining 3 combined")
```

	The boxplots below show that with respect to age, our data is right skewed, indicating 
	murders occurred at a higher frequency within the first three decades of our victims lives.
	This is supported with consistent medians, relatively close means and similar number of 
	outliers shown. The only standouts would be the Pacific, which has slightly more scattered 
	outliers, and the South which has the lowest percentage of murder victims below 30 ( 40% ).
	Based on the data, I suspect age brackets hardly impact murders among women below 30.
	

**Western Murders 1980-2014: N($\mu$ = 32.64, $\sigma$ = 17.42)**\newline
**Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.1515663$ making $\approx$ $44\%$ of murder victims years$\le$ **30**

```{r boxplot west, out.width="50%",echo=FALSE,results='hide'}
summary(west$`Victim Age`)
sd(west$`Victim Age`)
(30-32.64)/17.41812 
pnorm(-0.1515663)
normalPlot(bound = c(-Inf,-0.1515663))
ggplot(project_data, aes(y=`Victim Age`, x=femW))+geom_boxplot()+labs(title="Western Female gender vs. Remaining Regions",x="West & Female")
```
\
**Midwestern Murders 1980-2014: N($\mu$ = 32.6, $\sigma$ $\approx$ 17.34)**\newline
**Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.1499721$ making $\approx$ $44\%$ of murder victims years$\le$ **30**

```{r boxplot midwest, out.width="50%",echo=FALSE,results='hide'}
summary(midwest$`Victim Age`)
sd(midwest$`Victim Age`)
(30-32.6)/17.33656
pnorm(-0.1499721)
normalPlot(bound = c(-Inf,-0.1499721))
ggplot(project_data, aes(y=`Victim Age`, x=femMid))+geom_boxplot()+labs(title="MidWestern Female gender vs. Remaining Regions",x="MidWest & Female")
```

\clearpage

**Northeastern Murders 1980-2014: N($\mu$ = 32.99, $\sigma$ $\approx$ 17.56)**\newline
**Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.1702733$ making $\approx$ $43\%$ of murder victims years$\le$ **30**

```{r boxplot northeast, out.width="50%",echo=FALSE,results='hide'}
summary(northeast$`Victim Age`)
sd(northeast$`Victim Age`)
(30-32.99)/17.56
pnorm(-0.1702733)
normalPlot(bound = c(-Inf,-0.1702733))
ggplot(project_data, aes(y=`Victim Age`, x=femNE))+geom_boxplot()+labs(title="NorthEastern Female gender vs. Remaining Regions",x="NorthEast & Female")
```

**Southern Murders 1980-2014: N($\mu$ = 35, $\sigma$ $\approx$ 18.32)**\newline
**Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.2728698$ making $\approx$ $40\%$ of murder victims years$\le$ **30**

```{r boxplot south, out.width="50%",echo=FALSE,results='hide'}
summary(south$`Victim Age`)
sd(south$`Victim Age`)
(30-35)/ 18.32376
pnorm(-0.2728698)
normalPlot(bound = c(-Inf,-0.2728698))
ggplot(project_data, aes(y=`Victim Age`, x=femS))+geom_boxplot()+labs(title="Southern Female gender vs. Remaining Regions",x="Southern & Female")
```

**Pacific Murders 1980-2014: N($\mu$ = 33.81, $\sigma$ $\approx$ 17.12)**\newline
**Z = $\frac{x-\mu}{\sigma}$ Where $x$ $\le$ 30** = $-0.1168525$ making $\approx$ $45\%$ of murder victims years$\le$ **30**

```{r boxplot pacific, out.width="50%",echo=FALSE,results='hide'}
summary(pacific$`Victim Age`)
sd(pacific$`Victim Age`)
(30-32)/17.11559
pnorm(-0.1168525)
normalPlot(bound = c(-Inf,-0.1168525))
ggplot(project_data, aes(y=`Victim Age`, x=femP))+geom_boxplot()+labs(title="Pacific Female gender vs. Remaining Regions",x="Pacific & Female")

```

***

\clearpage

### Average Female Murder Count by Region

	Next I created a dataframe, categorizing the annual murder count of women by region
	for the years between 1980-2014. What I found appeared alarming at first. I seems the
	South has an overwhelming higher total female murder count:
\
\
\
	
\begin{tabular}{l  r  r  r  r  r  r}
                & \multicolumn{5}{c}{\textit{Female Murder Count by Region}} \\
\cline{2-6}
                & West  & Midwest  & Northeast & South & Pacific & Total/Ave \\
\hline
Mean            & 890.2     & 768     & 623.9714     & 1,785.943  & 25.37   & 818.69 \\
SD              & 160.84    & 163.9   & 150.17       & 240.88     & 7.23    & 144.6 \\
n               & 31,157    & 26,880  & 21,839       & 62,508     & 888     & 143,272 \\
\hline
\end{tabular}
\newline
\newline



```{r means_female_only, echo=FALSE, results='hide'}
female_only<-data.frame(Year = project_data$Year,Region = project_data$Region, Sex = project_data$`Victim Sex`)
glimpse(female_only)

```

```{r, echo=FALSE, results='hide'}
female_only<-female_only[female_only$Sex=='Female',]
glimpse(female_only)

```


```{r, echo=FALSE, results='hide'}
female_only<-female_only%>%count(Sex,Region,Year)
glimpse(female_only)
```




```{r,echo=FALSE, results='hide'}
female_only%>%filter(Region=="West")%>%summarize(mean(n))
female_only%>%filter(Region=="Midwest")%>%summarize(mean(n))
female_only%>%filter(Region=="Northeast")%>%summarize(mean(n))
female_only%>%filter(Region=="South")%>%summarize(mean(n))
female_only%>%filter(Region=="Pacific")%>%summarize(mean(n))
```




```{r sd_female_only, echo=FALSE, results='hide'}
female_only%>%filter(Region=="West")%>%summarize(sd(n))
female_only%>%filter(Region=="Midwest")%>%summarize(sd(n))
female_only%>%filter(Region=="Northeast")%>%summarize(sd(n))
female_only%>%filter(Region=="South")%>%summarize(sd(n))
female_only%>%filter(Region=="Pacific")%>%summarize(sd(n))
```

```{r sum_female_only, echo=FALSE, results='hide'}
female_only%>%filter(Region=="West")%>%summarize(sum(n))

```

```{r, echo=FALSE, results='hide'}
female_only%>%filter(Region=="Midwest")%>%summarize(sum(n))

```

```{r, echo=FALSE, results='hide'}
female_only%>%filter(Region=="Northeast")%>%summarize(sum(n))

```

```{r, echo=FALSE, results='hide'}
female_only%>%filter(Region=="South")%>%summarize(sum(n))

```

```{r}
female_only%>%filter(Region=="Pacific")%>%summarize(sum(n))
```


```{r, echo=FALSE, results='hide'}
mean(female_only$n)
sd(female_only$n)
sum(female_only$n)
project_data %>% filter(`Victim Sex` == "Female" & `Victim Age`< 31) %>% count()
```

```{r, echo=FALSE, out.height="100%"}

ggplot(data = female_only, aes(x = Year, y = n, color=Region, group = Region, )) +
  geom_line() +
	geom_point()+
	ggtitle("Annual Female Murder Count by Region")+
	ylab("Muder Count")
```

```{r echo=FALSE, warning=FALSE}
paste0("Data Frame: female_only<-data.frame(Year = project_data$Year,R...")
paste0( " [female_only$Sex!='Male',] ...count(Sex,Region,Year)  ")
```
\clearpage

## Part 4 - Inference

	The data above shows similar behavior with respect to mean age of murder count.
	It is also apparent that gender, impacts murder counts when comparing Males to
	Females. Now we will consider the impact Regions have on murder counts of cases
	of the same gender. I will infer based on age, by categorizing Females into 2
	groups, those above 30 years of age and those below. The decision to split at 
	30 years, comes from looking at the National average 33.56, and my largest group
	set for age, the 20-30 bracket. The 2 regions I will use are the highest and 
	lowest for my Annual "Female Murder Count by Region" line graph, the South and 
	the Pacific.
	

### South Inference

**Inference Conditions**

* Independence - Murder counts are assumed independent with 62,508 observations$\checkmark$
* Success-failure Conditions - $Count_{under\ 30}$=27,982 & $Count_{over\ 30}$=34,526 $\therefore n \ge 30$ $\checkmark$

**Hypothesis statement**\newline
$H_0$ = $South_{30yrs\ OR\ less} = 0.\ No\ Difference\ in\ average\ below\ or\ above\ 30\ years\ for\ women$\newline
$H_A$ = $South_{30yrs\ OR\ less} \ne 0.\ There\ is\ a\ difference$\newline


```{r south inference female batch, echo=TRUE, results='hide'}

south<-south[south$`Victim Sex`=='Female',]
glimpse(south)
#count(south)
south<-south%>%count(`Victim Sex`,`Victim Age`,Region,Year)

```


```{r new var under30 south, echo=TRUE}
#New Variable Created 'under30'
south <- south %>% 
  mutate(under30 = ifelse(south$`Victim Age` < 31, "yes", "no"))
```

```{r, echo=FALSE,results='hide'}
south%>%filter(under30=="yes")%>%summarize(sum(n))
south%>%filter(under30=="no")%>%summarize(sum(n))
```


```{r by-means south,include=FALSE, echo=FALSE, results='hide'}
south %>%
  group_by(under30) %>%
  summarise(mean_n = mean(n, na.rm = TRUE))
```

```{r inf-weight-habit-ht-south-initial, tidy=FALSE, warning = FALSE,echo=TRUE}
obs_diff_south <- south %>%
  specify(n ~ under30) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```

Observation stat of 10.39966 generated & `hypothesize` function used to set null hypothesis as a test for independence. We have 0 null permutations.

```{r inf-weight-habit-ht-null-south, tidy=FALSE, warning = FALSE, echo=TRUE}
null_dist_south <- south %>%
  specify(n ~ under30) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```

```{r, include=FALSE}
sum(null_dist_south$stat > 10.39966)
```

```{r south histo,warning=FALSE, echo=FALSE, message=FALSE, out.width="50%"}
ggplot(data = null_dist_south, aes(x = stat)) +
  geom_histogram()

ggplot(south, aes(x = n, colour = under30)) + geom_density()
```

```{r inf-weight-habit-ht-pvalue, echo=FALSE,warning=FALSE}
null_dist_south %>%
  get_p_value(obs_stat = obs_diff_south, direction = "two_sided")
```

\clearpage

```{r inference function,echo=TRUE, results='markup',fig.height=3}
inference(y = south$n, x = south$under30,est = "mean", type = "ci", method = "theoretical")
```

\clearpage

### Pacific Inference

**Inference Conditions**

* Independence - Murder counts are assumed independent with 888 observations$\checkmark$
* Success-failure Conditions - $Count_{under\ 30}$=421 & $Count_{over\ 30}$=467 $\therefore n \ge 30$ $\checkmark$

**Hypothesis statement**\newline
$H_0$ = $Pacific_{30yrs\ OR\ less} = 0.\ No\ Difference\ in\ average\ below\ or\ above\ 30\ years\ for\ women$\newline
$H_A$ = $Pacific_{30yrs\ OR\ less} \ne 0.\ There\ is\ a\ difference$\newline


```{r pacific inference female batch, echo=FALSE, results='hide'}

pacific<-pacific[pacific$`Victim Sex`=='Female',]
count(pacific)
pacific<-pacific%>%count(`Victim Sex`,`Victim Age`,Region,Year)

```


```{r new var under30 Pacific, echo=TRUE}
#New Variable Created 'under30'
pacific <- pacific %>% 
  mutate(under30 = ifelse(pacific$`Victim Age` < 31, "yes", "no"))
```


```{r, echo=FALSE,results='hide'}
pacific%>%filter(under30=="yes")%>%summarize(sum(n))
pacific%>%filter(under30=="no")%>%summarize(sum(n))
```


```{r by-means Pacific,include=FALSE, echo=FALSE, results='hide',warning=FALSE}
pacific %>%
  group_by(under30) %>%
  summarise(mean_n = mean(n, na.rm = TRUE))
```

```{r inf-weight-habit-ht-initial, tidy=FALSE, warning = FALSE,echo=FALSE}
obs_diff_pacific <- pacific %>%
  specify(n ~ under30) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```

Observation stat of 0.1601648 generated & `hypothesize` function used to set null hypothesis as a test for independence. We have 0 null permutations.

```{r inf-weight-habit-ht-null, tidy=FALSE, warning = FALSE, echo=FALSE}
null_dist_pacific <- pacific %>%
  specify(n ~ under30) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```

```{r, include=FALSE}
sum(null_dist_pacific$stat > 0.1601648)
```

```{r Pacific histo,warning=FALSE, echo=FALSE, message=FALSE, out.width="50%"}
ggplot(data = null_dist_pacific, aes(x = stat)) +
  geom_histogram()

ggplot(pacific, aes(x = n, colour = under30)) + geom_density()
```

```{r inf-weight-habit-ht-pvalue-pacific, echo=FALSE,warning=FALSE}
null_dist_pacific %>%
  get_p_value(obs_stat = obs_diff_pacific, direction = "two_sided")
```

\clearpage

```{r inference function pacific,echo=TRUE, results='markup',fig.height=3}
inference(y = pacific$n, x = pacific$under30,est = "mean", type = "ci", method = "theoretical")
```



## Part 5 - Conclusion

My conclusion is based on the following facts:
	
* Mean age on a national and a regional level hardly differs
* Region does not impact murder count of a gender by age
* Gender does impact murder count, but only when referencing male vs. female

Noting the higher murder count for the South vs. the Pacific, considering the Pacific is easily much smaller
being composed of only 2 states, and the fact that both follow national trends indicating peak numbers
occur for victims 20-30, I would say the difference is most likely because of population size. If I were to
do a comparison of population of these regions during these years, I'd imagine the murder counts would be
comparable. Therefore, my limitation with this data is not having the population size for each states, per year, between 1980-2014. Regardless the data does not highlight any specific region as having an abnormal murder count in my opinion. I can support that age is the biggest factor among those mentions.


## References:

### Homicide Reports, 1980-2014

*Project
https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv*


