---
title: Test data project
author: Gabriel Campos
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
paste0(" Attaching packages: DATA606,infer,dplyr,VennDiagram,scales,data.table,readr ")
```

```{r echo=FALSE, warning=FALSE, message=FALSE,include=FALSE}
library(DATA606)
library(infer)
library(dplyr)
library(VennDiagram)
#library(scales)
library(data.table)
library (readr)
```



### Load Data

```{r vecotr-region, echo=FALSE, message=FALSE, warning=FALSE}
project_data<- read_csv("D:/Bulk Data/database.csv")
project_data$Region<-
	ifelse(project_data$State=="Arizona" |
		   	project_data$State=="California"| project_data$State=="Colorado"| project_data$State=="Idaho"|
		   	project_data$State=="Montana"| project_data$State=="Nevada"| project_data$State=="New Mexico"|
		   	project_data$State=="Oregon"| project_data$State=="Utah"| project_data$State=="Washington"|
		   	project_data$State=="Wyoming","West",
		   ifelse(project_data$State=="Illinois"|
		   	   	project_data$State == "Indiana"| project_data$State == "Iowa"| project_data$State == "Kansas"|
		   	   	project_data$State == "Michigan"| project_data$State == "Minnesota"|
		   	   	project_data$State == "Missouri"| project_data$State == "Nebraska"|
		   	   	project_data$State == "North Dakota"| project_data$State == "Ohio"|
		   	   	project_data$State == "South Dakota"| project_data$State == "Wisconsin","Midwest",
		   	   ifelse(project_data$State == "Connecticut"|
		   	   	   	project_data$State == "Delaware"| project_data$State =="Maine"|
		   	   	   	project_data$State =="Massachusetts"| project_data$State =="New Hampshire"|
		   	   	   	project_data$State =="New Jersey"| project_data$State == "New York"|
		   	   	   	project_data$State =="Pennsylvania"| project_data$State == "Rhodes Island"|
		   	   	   	project_data$State=="Vermont","Northeast",
		   	   	   ifelse(project_data$State == "Alabama"| project_data$State == "Arkansas"|
		   	   	   	   	project_data$State == "Delaware"| project_data$State =="District of Columbia"|
		   	   	   	   	project_data$State == "Florida" | project_data$State == "Georgia"|
		   	   	   	   	project_data$State == "Kentucky"| project_data$State ==	"Louisiana"|
		   	   	   	   	project_data$State == "Maryland"| project_data$State == "Mississippi"|
		   	   	   	   	project_data$State == "North Carolina"| project_data$State == "Oklahoma"|
		   	   	   	   	project_data$State == "South Carolina"| project_data$State == "Tennessee"|
		   	   	   	   	project_data$State == "Texas"| project_data$State == "Virginia"|
		   	   	   	   	project_data$State == "West Virginia","South",
		   	   	   	   ifelse(project_data$State =="Alaska"| project_data$State == "Hawaii", "Pacific",
		   	   	   	   	   "Other"))
		   	   	   )
		   	   )
		   )

project_data$`Victim AgeGroup`<- cut(project_data$`Victim Age`,
									 breaks=c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100),
									 right = FALSE)

project_data %>% filter(Region == "Other") %>% glimpse()
```



```{r, echo=FALSE, warning=FALSE, message=FALSE}
#make region
west <- project_data %>%
  filter(State == "Arizona"| State == "California"| State =="Colorado"
  	   | State =="Idaho"| State =="Montana"| State =="Nevada"
  	   | State == "New Mexico"| State =="Oregon" | State == "Utah"
  	   | State =="Washington"| State =="Wyoming")
#Filter 998
west<-west[!(west$`Victim Age`==998) & west$`Victim Sex`!='Unknown',]

paste0("West")
west %>% filter(`Victim Sex` == "Female") %>% glimpse()

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#make region
midwest <- project_data %>%
  filter(State == "Illinois"| State == "Indiana"| State =="Iowa"
  	   | State =="Kansas"| State =="Michigan"| State =="Minnesota"
  	   | State == "Missouri"| State =="Nebraska"| State == "North Dakota"
  	   | State =="Ohio"| State =="South Dakota"| State =="Wisconsin")
#Filter 998
midwest<-midwest[!(midwest$`Victim Age`==998) & midwest$`Victim Sex`!='Unknown',]

paste0("midwest")
midwest %>% filter(`Victim Sex` == "Female") %>% glimpse()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#make region
northeast <- project_data %>%
  filter(State == "Connecticut"| State == "Delaware"| State =="Maine"
  	   | State =="Massachusetts"| State =="New Hampshire"
  	   | State =="New Jersey"| State == "New York"| State =="Pennsylvania"
  	   | State == "Rhodes Island"| State =="Vermont")
#Filter 998
northeast<-northeast[!(northeast$`Victim Age`==998) & northeast$`Victim Sex`!='Unknown',]

paste0("northeast")
northeast %>% filter(`Victim Sex` == "Female") %>% glimpse()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#make region
south <- project_data %>%
  filter(State == "Alabama"| State == "Arkansas"| State == "Delaware"
  	   | State =="District of Columbia"| State =="Florida" 
  	   | State =="Georgia"| State =="Kentucky"
  	   | State =="Louisiana"| State == "Maryland"|State == "Mississippi"
  	   | State =="North Carolina"| State == "Oklahoma"
  	   | State =="South Carolina"| State =="Tennessee"
  	   | State =="Texas"| State =="Virginia"| State =="West Virginia")
#Filter 998
south<-south[!(south$`Victim Age`==998) & south$`Victim Sex`!='Unknown',]

paste0("south")
south %>% filter(`Victim Sex` == "Female") %>% glimpse()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#make region
pacific<- project_data %>%
  filter(State == "Alaska"| State == "Hawaii")
#Filter 998
pacific<-pacific[!(pacific$`Victim Age`==998) & pacific$`Victim Sex`!='Unknown',]

paste0("pacific")
pacific %>% filter(`Victim Sex` == "Female") %>% glimpse()
```

```{r, echo=FALSE}
#count<-project_data$`Victim Age`
#table(count)
project_data<-project_data[!(project_data$`Victim Age`==998) & project_data$`Victim Sex`!='Unknown',]
project_data %>% filter(`Victim Sex` == "Female" & `Victim Age`< 31) %>% glimpse()
glimpse(project_data)

```


```{r, echo=FALSE}
summary(project_data$`Victim Age`)
sd(project_data$`Victim Age`)
```


```{r create new var full, echo=FALSE}
#WEST
project_data <- project_data %>% 
  mutate(femW = ifelse(project_data$Region=="West" & project_data$`Victim Sex`=="Female", "yes", "no"))


#MIDWEST
project_data <- project_data %>% 
  mutate(femMid = ifelse(project_data$Region=="Midwest" & project_data$`Victim Sex`=="Female", "yes", "no"))

#NORTHEAST
project_data <- project_data %>% 
  mutate( femNE= ifelse(project_data$Region=="Northeast" & project_data$`Victim Sex`=="Female", "yes", "no"))

#SOUTH
project_data <- project_data %>% 
  mutate( femS= ifelse(project_data$Region=="South" & project_data$`Victim Sex`=="Female", "yes", "no"))

#PACIFIC
project_data <- project_data %>% 
  mutate( femP= ifelse(project_data$Region=="Pacific" & project_data$`Victim Sex`=="Female", "yes", "no"))
```

```{r boxplot west, out.width="50%",echo=FALSE,results='show'}
summary(west$`Victim Age`)
sd(west$`Victim Age`)
#(30-32.64)/17.41812 
#pnorm(-0.1515663)
#normalPlot(bound = c(-Inf,-0.1515663))
#ggplot(project_data, aes(y=`Victim Age`, x=femW))+geom_boxplot()+labs(title="Western Female gender vs. Remaining Regions",x="West & Female")
```


```{r boxplot midwest, out.width="50%",echo=FALSE,results='hide'}
summary(midwest$`Victim Age`)
sd(midwest$`Victim Age`)
#(30-32.6)/17.33656
#pnorm(-0.1499721)
#normalPlot(bound = c(-Inf,-0.1499721))
#ggplot(project_data, aes(y=`Victim Age`, x=femMid))+geom_boxplot()+labs(title="MidWestern Female gender vs. Remaining Regions",x="MidWest & Female")
```



```{r boxplot northeast, out.width="50%",echo=FALSE,results='hide'}
summary(northeast$`Victim Age`)
sd(northeast$`Victim Age`)
#(30-32.99)/17.56
#pnorm(-0.1702733)
#normalPlot(bound = c(-Inf,-0.1702733))
#ggplot(project_data, aes(y=`Victim Age`, x=femNE))+geom_boxplot()+labs(title="NorthEastern Female gender vs. Remaining Regions",x="NorthEast & Female")
```



```{r boxplot south, out.width="50%",echo=FALSE,results='hide'}
summary(south$`Victim Age`)
sd(south$`Victim Age`)
#(30-35)/ 18.32376
#pnorm(-0.2728698)
#normalPlot(bound = c(-Inf,-0.2728698))
#ggplot(project_data, aes(y=`Victim Age`, x=femS))+geom_boxplot()+labs(title="Southern Female gender vs. Remaining Regions",x="Southern & Female")
```


```{r boxplot pacific, out.width="50%",echo=FALSE,results='hide'}
summary(pacific$`Victim Age`)
sd(pacific$`Victim Age`)
#(30-32)/17.11559
#pnorm(-0.1168525)
#normalPlot(bound = c(-Inf,-0.1168525))
#ggplot(project_data, aes(y=`Victim Age`, x=femP))+geom_boxplot()+labs(title="Pacific Female gender vs. Remaining Regions",x="Pacific & Female")

```


\
\
\
	
\begin{tabular}{l  r  r  r  r  r  r}
                & \multicolumn{5}{c}{\textit{Female Murder Count by Region}} \\
\cline{2-6}
                & West  & Midwest  & Northeast & South & Pacific & Total \\
\hline
Mean            & 890.2     & 768     & 623.65       & 1,786.25   & 25.37   & 818.6971 \\
SD              & 160.84    & 163.9   & 149.09       & 241.81     & 7.23    & 591.7687 \\
n               & 31,157    & 26,880  & 21,828        & 62,519     & 888     & 143,272 \\
\hline
\end{tabular}
\newline
\newline



```{r means_female_only, echo=FALSE, results='hide'}
female_only<-data.frame(Year = project_data$Year,Region = project_data$Region, Sex = project_data$`Victim Sex`, Age =project_data$`Victim Age`)
glimpse(female_only)
female_only %>% filter(Sex=='Female' & Age<31) %>% glimpse()

```

```{r, echo=FALSE, results='hide'}
female_only<-female_only[female_only$Sex=='Female',]
glimpse(female_only)
```
```{r}
female_only %>% filter(Age<31) %>% glimpse()
```

```{r}
female_only %>% filter(Region == 'South') %>% glimpse()
```


```{r, echo=FALSE, results='hide'}
female_only<-female_only%>%count(Sex,Region,Year)
glimpse(female_only)
```




```{r,echo=FALSE, results='hide'}
female_only%>%filter(Region=="West")%>%summarize(mean(n))
female_only%>%filter(Region=="Midwest")%>%summarize(mean(n))
female_only%>%filter(Region=="Northeast")%>%summarize(mean(n))
female_only%>%filter(Region=="South")%>%summarize(mean(n))
female_only%>%filter(Region=="Pacific")%>%summarize(mean(n))
```




```{r sd_female_only, echo=FALSE, results='hide'}
female_only%>%filter(Region=="West")%>%summarize(sd(n))
female_only%>%filter(Region=="Midwest")%>%summarize(sd(n))
female_only%>%filter(Region=="Northeast")%>%summarize(sd(n))
female_only%>%filter(Region=="South")%>%summarize(sd(n))
female_only%>%filter(Region=="Pacific")%>%summarize(sd(n))
```

```{r sum_female_only, echo=FALSE, results='hide'}
female_only%>%filter(Region=="West")%>%summarize(sum(n))

```

```{r, echo=FALSE, results='hide'}
female_only%>%filter(Region=="Midwest")%>%summarize(sum(n))

```

```{r, echo=FALSE, results='hide'}
female_only%>%filter(Region=="Northeast")%>%summarize(sum(n))

```

```{r, echo=FALSE, results='hide'}
female_only%>%filter(Region=="South")%>%summarize(sum(n))

```

```{r}
female_only%>%filter(Region=="Pacific")%>%summarize(sum(n))
```


```{r, echo=FALSE, results='hide'}
mean(female_only$n)
sd(female_only$n)
sum(female_only$n)
project_data %>% filter(`Victim Sex` == "Female" & `Victim Age`< 31) %>% count()
```

```{r, echo=FALSE, out.height="100%"}

#ggplot(data = female_only, aes(x = Year, y = n, color=Region, group = Region, )) +
#  geom_line() +
#	geom_point()+
#	ggtitle("Annual Female Murder Count by Region")+
#	ylab("Muder Count")
```

## Part 4 - Inference



### South Inference
```
**Inference Conditions**

* Independence - Murder counts are assumed independent with 62,519 observations$\checkmark$
* Success-failure Conditions - $Count_{under\ 30}$=1079 & $Count_{over\ 30}$=2212 $\therefore n \ge 30$ $\checkmark$

**Hypothesis statement**\newline
$H_0$ = $South_{30yrs\ OR\ less} = 0.\ No\ Difference\ in\ average\ below\ or\ above\ 30\ years\ for\ women$\newline
$H_A$ = $South_{30yrs\ OR\ less} \ne 0.\ There\ is\ a\ difference$\newline


```{r south inference female batch, echo=FALSE, results='hide'}

#south<-south[south$`Victim Sex`=='Female',]
#glimpse(south)
#count(south)
#south<-south%>%count(`Victim Sex`,`Victim Age`,Region,Year)

```


```{r new var under30 south, echo=TRUE}
#New Variable Created 'under30'
#south <- south %>% 
#  mutate(under30 = ifelse(south$`Victim Age` < 31, "yes", "no"))
```

```{r by-means south,include=FALSE, echo=FALSE, results='hide'}
#south %>%
#  group_by(under30) %>%
#  summarise(mean_n = mean(n, na.rm = TRUE))
```

```{r inf-weight-habit-ht-south-initial, tidy=FALSE, warning = FALSE,echo=FALSE}
#obs_diff_south <- south %>%
#  specify(n ~ under30) %>%
#  calculate(stat = "diff in means", order = c("yes", "no"))
```

Observation stat of 8.593917 generated & `hypothesize` function used to set null hypothesis as a test for independence. We have 0 null permutations.

```{r inf-weight-habit-ht-null-south, tidy=FALSE, warning = FALSE, echo=FALSE}
#null_dist_south <- south %>%
#  specify(n ~ under30) %>%
#  hypothesize(null = "independence") %>%
#  generate(reps = 1000, type = "permute") %>%
#  calculate(stat = "diff in means", order = c("yes", "no"))
```

```{r, include=FALSE}
#sum(null_dist_south$stat > 8.59)
```

```{r south histo,warning=FALSE, echo=FALSE, message=FALSE, out.width="50%"}
#ggplot(data = null_dist_south, aes(x = stat)) +
#  geom_histogram()

#ggplot(south, aes(x = n, colour = under30)) + geom_density()
```

```{r inf-weight-habit-ht-pvalue, echo=FALSE,warning=FALSE}
#null_dist_south %>%
#  get_p_value(obs_stat = obs_diff_south, direction = "two_sided")
```

\clearpage

```{r inference function,echo=TRUE, results='markup',fig.height=3}
#inference(y = south$n, x = south$under30,est = "mean", type = "ci", method = "theoretical")
```

\clearpage

### Pacific Inference

**Inference Conditions**

* Independence - Murder counts are assumed independent with 888 observations$\checkmark$
* Success-failure Conditions - $Count_{under\ 30}$=308 & $Count_{over\ 30}$=387 $\therefore n \ge 30$ $\checkmark$

**Hypothesis statement**\newline
$H_0$ = $Pacific_{30yrs\ OR\ less} = 0.\ No\ Difference\ in\ average\ below\ or\ above\ 30\ years\ for\ women$\newline
$H_A$ = $Pacific_{30yrs\ OR\ less} \ne 0.\ There\ is\ a\ difference$\newline


```{r pacific inference female batch, echo=FALSE, results='hide'}

#pacific<-pacific[pacific$`Victim Sex`=='Female',]
#count(pacific)
#pacific<-pacific%>%count(`Victim Sex`,`Victim Age`,Region,Year)

```


```{r new var under30 Pacific, echo=TRUE}
#New Variable Created 'under30'
#pacific <- pacific %>% 
# mutate(under30 = ifelse(pacific$`Victim Age` < 31, "yes", "no"))
```

```{r by-means Pacific,include=FALSE, echo=FALSE, results='hide',warning=FALSE}
#pacific %>%
#  group_by(under30) %>%
#  summarise(mean_n = mean(n, na.rm = TRUE))
```

```{r inf-weight-habit-ht-initial, tidy=FALSE, warning = FALSE,echo=FALSE}
#obs_diff_pacific <- pacific %>%
#  specify(n ~ under30) %>%
#  calculate(stat = "diff in means", order = c("yes", "no"))
```

Observation stat of 0.1601648 generated & `hypothesize` function used to set null hypothesis as a test for independence. We have 0 null permutations.

```{r inf-weight-habit-ht-null, tidy=FALSE, warning = FALSE, echo=FALSE}
#null_dist_pacific <- pacific %>%
#  specify(n ~ under30) %>%
#  hypothesize(null = "independence") %>%
#  generate(reps = 1000, type = "permute") %>%
#  calculate(stat = "diff in means", order = c("yes", "no"))
```

```{r, include=FALSE}
#sum(null_dist_pacific$stat > 0.1601648)
```

```{r Pacific histo,warning=FALSE, echo=FALSE, message=FALSE, out.width="50%"}
#ggplot(data = null_dist_pacific, aes(x = stat)) +
#  geom_histogram()

#ggplot(pacific, aes(x = n, colour = under30)) + geom_density()
```

```{r inf-weight-habit-ht-pvalue-pacific, echo=FALSE,warning=FALSE}
#null_dist_pacific %>%
#  get_p_value(obs_stat = obs_diff_pacific, direction = "two_sided")
```

\clearpage

```{r inference function pacific,echo=TRUE, results='markup',fig.height=3}
#inference(y = pacific$n, x = pacific$under30,est = "mean", type = "ci", method = "theoretical")
```



## Part 5 - Conclusion

My conclusion is based on the following facts:
	
* Mean age on a national and a regional level hardly differs
* Region does not impact murder count of a gender by age
* Gender does impact murder count, but only when referencing male vs. female

Noting the higher murder count for the South vs. the Pacific, considering the Pacific is easily much smaller
being composed of only 2 states, and the fact that both follow national trends indicating peak numbers
occur for victims 20-30, I would say the difference is most likely because of population size. If I were to
do a comparison of population of these regions during these years, I'd imagine the murder counts would be
comparable. Therefore, my limitation with this data is not having the population size for each states, per year, between 1980-2014. Regardless the data does not highlight any specific region as having an abnormal murder count in my opinion. I can support that age is the biggest factor among those mentions.


## References:

### Homicide Reports, 1980-2014

*Project
https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv*


```
