---
title: "Lab2: Intro to Data"
author: "Gabriel Campos"
output: 
 pdf_document:
        extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
data("nycflights")
```

[**Basic R Markdown with an OpenIntro Lab**](https://www.youtube.com/watch?v=Pdc368lS2hk)
<iframe width="560" height="315" src="https://www.youtube.com/embed/Pdc368lS2hk" frameborder="0" allowfullscreen></iframe>

### Lab report
To record your analysis in a reproducible format, you can adapt the general Lab 
Report template from the **openintro** package. Watch the video above to learn how.



```{r hist-dep-delay}
names(nycflights)
```

### Exercise 1
1.  Look carefully at these three histograms. How do they compare? Are features
revealed in one that are obscured in another?
```{r plot-hist-nycflights}
ggplot(data=nycflights, aes(x = dep_delay)) +
	geom_histogram()
```
```{r hist-dep-delay-bin-15}
ggplot(data = nycflights, aes(x = dep_delay))+
	geom_histogram(binwidth = 15)
```
```{r hist-dep_delay-bin-150}
ggplot(data = nycflights, aes(x = dep_delay)) +
	geom_histogram(binwidth = 150)
```


### __Exercise 1 Answer:__
**i)** binwidth = 15 shows departure delays before the peak amount and is symmetrical\newline\
**ii)** binwidth = 150 shows a larger count for number of delay flights. Approximately 100 more and is left skewed.

\

### Exercise 2
2.  Create a new data frame that includes flights headed to SFO in February, 
    and save this data frame as `sfo_feb_flights`. How many flights 
    meet these criteria?
```{r sfo_feb_flights}
sfo_feb_flights <- nycflights %>%
	filter(dest == "SFO", month == 2)
```

```{r glimpse-sfo_feb_flights}
glimpse(sfo_feb_flights)
```

### __Exercise 2 Answer:__
68 flights meet these criteria.\newline

### Exercise 3
3.  Describe the distribution of the **arrival** delays of these flights using a 
    histogram and appropriate summary statistics. **Hint:** The summary 
    statistics you use should depend on the shape of the distribution.
    
```{r ggplot-hist-sfo_feb_arr_delay}
ggplot(data = sfo_feb_flights, aes(x = arr_delay))+
	geom_histogram(binwidth = 10)
```
```{r hist-plot-sfo_feb_arr_delay}
### Alternative method based off of https://www.statmethods.net/graphs/density.html
hist(sfo_feb_flights$arr_delay)
```

*Note: Histogram is symmetric and unimodal with one potential outlier.*\newline \newline

### __Exercise 3 Answer:__
For the bell curve created by the histogram, the appropriate summary statistics are the numerical stats mean, media, interquartile range, standard deviation and number of values "n".

```{r summarise-sfo_feb_num_stats}
sfo_feb_flights %>%
  summarise(mean_ad   = mean(arr_delay), 
            median_ad = median(arr_delay),
  			iqr_ad = IQR(arr_delay),
  			sd_ad = sd(arr_delay),
            n_ad = n())
```

### Exercise 4

Calculate the median and interquartile range for `arr_delay`s of flights in the `sfo_feb_flights` data frame, grouped by carrier. Which carrier has the most variable arrival delays?

```{r summarise-group-sfo_feb_num_stats}
sfo_feb_flights %>%
	group_by(carrier) %>%
	summarise(median_ad = median(arr_delay), iqr_ad = IQR(arr_delay))
```

### __Exercise 4 Answer:__

United Airlines Inc. (UA) and Delta Airlines Inc. (DL) have the highest interquartile range. IQR is a measure of variability. Hence the answer is UA and DL

### Exercise 5

Suppose you really dislike departure delays and you want to schedule 
your travel in a month that minimizes your potential departure delay leaving 
NYC. One option is to choose the month with the lowest mean departure delay.
Another option is to choose the month with the lowest median departure delay. 
What are the pros and cons of these two choices?
    
    
```{r month_means}
nycflights %>%
	group_by(month) %>%
	summarise(mean_dd = mean(dep_delay), median(dep_delay))
```


### __Exercise 5 Answer:__

Mean is the optimal choice. An average gives a calculated value of a the potential delay for the month. The median is uncalculated, robust and less sensitive to outliers in the data set for the month (e.g. 1/2013 saw at least one triple digit delay) hence is not the best choice.

**On time departure rate for NYC airports**

Suppose you will be flying out of NYC and want to know which of the 
three major NYC airports has the best on time departure rate of departing flights. 
Also supposed that for you, a flight that is delayed for less than 5 minutes is 
basically "on time."" You consider any flight delayed for 5 minutes of more to be 
"delayed".

In order to determine which airport has the best on time departure rate, 
you can 

- first classify each flight as "on time" or "delayed",
- then group flights by origin airport,
- then calculate on time departure rates for each origin airport,
- and finally arrange the airports in descending order for on time departure
percentage.

Let's start with classifying each flight as "on time" or "delayed" by
creating a new variable with the `mutate` function.

```{r dep-type}
nycflights <- nycflights %>%
  mutate(dep_type = ifelse(dep_delay < 5, "on time", "delayed"))
```

The first argument in the `mutate` function is the name of the new variable
we want to create, in this case `dep_type`. Then if `dep_delay < 5`, we classify 
the flight as `"on time"` and `"delayed"` if not, i.e. if the flight is delayed 
for 5 or more minutes.

Note that we are also overwriting the `nycflights` data frame with the new 
version of this data frame that includes the new `dep_type` variable.

We can handle all of the remaining steps in one code chunk:

```{r ot-dep-rate}
nycflights %>%
  group_by(origin) %>%
  summarise(ot_dep_rate = sum(dep_type == "on time") / n()) %>%
  arrange(desc(ot_dep_rate))
```

### Exercise 6
If you were selecting an airport simply based on on time departure 
percentage, which NYC airport would you choose to fly out of?
    
### __Exercise 6 Answer:__
I would chose LGA or LaGuardia which has a 72.3% on time arrival percentage

### Exercise 7
Mutate the data frame so that it includes a new variable that contains the 
average speed, `avg_speed` traveled by the plane for each flight (in mph).
**Hint:** Average speed can be calculated as distance divided by
number of hours of travel, and note that `air_time` is given in minutes.

### __Exercise 7 Answer:__

```{r calc-avg_speed-view}
# This chunk was in order to view code before storing on table
#nycflights%>%
#	mutate(avg_speed = (distance/(air_time/60)))
```


```{r calc-avg_speed}
nycflights<- nycflights%>%
	mutate(avg_speed = (distance/(air_time/60)))
```

```{r glimpse-table-nycflights}
# to verify storage
glimpse(nycflights)
```
**Note the last variable is avg_speed**

### Exercise 8
Make a scatterplot of `avg_speed` vs. `distance`. Describe the relationship
between average speed and distance.
**Hint:** Use `geom_point()`.

```{r scat_ggplot-nycflights-avg_spd_dist}
ggplot(data = nycflights, mapping = aes(x = distance,y = avg_speed)) + geom_point()
```

### __Exercise 8 Answers__
There is a positive association with avg_speed and distance

### Exercise 9
Replicate the following plot. **Hint:** The data frame plotted only
contains flights from American Airlines, Delta Airlines, and United
Airlines, and the points are `color`ed by `carrier`. Once you replicate
the plot, determine (roughly) what the cutoff point is for departure
delays where you can still expect to get to your destination on time.

### __Exercise 9 Answer__
First create the data frame with the appropriate data
```{r select-ex9_data}
ex9_data<-nycflights%>%
	filter(carrier=="AA"|carrier=="DL"|carrier=="UA")
```

then create the scatter plot with color configuration using dep_delay and arr_delay based on graph provided.
```{r sctr_ggplot-ex9_data}
ggplot(data  = ex9_data, mapping = aes(x = dep_delay, y = arr_delay, color = carrier))+ geom_point()
```
```{r sctr_gg_box_plot-ex9_data}
ggplot(data = ex9_data, aes(group = 1, x = dep_delay, y = arr_delay, color = carrier))+ geom_point() + ylim(-100,0) +xlim(-100, 65)
```

```{r calc-max_y_where_x_zero}
ex9_data%>%
	summarise(max_dep = max(dep_delay[arr_delay==0]))
```
### __Exercise 9 Answers Part II__
\newline

Assessing the latest a person can depart and still arrive on time can be viewed as largest departure delay value where the arrival delay value equals 0. Therefore I set the app_delay limit value between -100 to 0. From there I decreased the x-limit value down to 65 (as decreasing any more would make my largest plot point disappear).\newline
\newline
\newline
To check my result I also searched the largest departure delay where arrival delay equals 0 using the max function.\newline
\newline \newline
I found that (roughly) the cutoff point for departure is between **40-65** minutes.

...

