---
title: DATA 606 Data Project Proposal
author: Gabriel Campos
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
library(tidyverse)
library (readr)
urlfile="https://raw.githubusercontent.com/fivethirtyeight/data/master/bad-drivers/bad-drivers.csv"
project_data<-read_csv(url(urlfile))
```

**Elements in project data are: **

```{r, echo=FALSE}
glimpse(project_data)
```
\
**All 51 states make up the data set**
\
```{r, echo=FALSE}
str(project_data$State)
```
\
**5 subsets based on `project_data` are made from project data based on region**
\

```{r, include=FALSE}
west <- project_data %>%
  filter(State == "Arizona"| State == "California"| State =="Colorado"| State =="Idaho"| State =="Montana"
  	   | State =="Nevada"| State == "New Mexico"| State =="Oregon" | State == "Utah"| State =="Washington"
  	   | State =="Wyoming")
midwest <- project_data %>%
  filter(State == "Illinois"| State == "Indiana"| State =="Iowa"| State =="Kansas"| State =="Michigan"
  	   | State =="Minnesota"| State == "Missouri"| State =="Nebraska" | State == "North Dakota"| State =="Ohio"
  	   | State =="South Dakota"| State =="Wisconsin")
northeast <- project_data %>%
  filter(State == "Connecticut"| State == "Delaware"| State =="Maine"| State =="Massachusetts"
  	   | State =="New Hampshire"| State =="New Jersey"| State == "New York"| State =="Pennsylvania"
  	   | State == "Rhode Island"| State =="Vermont")
south <- project_data %>%
  filter(State == "Alabama"| State == "Arkansas"| State =="Florida"| State =="Georgia"| State =="Kentucky"
  	   | State =="Louisiana"| State == "Mississippi"| State =="North Carolina"| State == "Oklahoma"
  	   | State =="South Carolina"| State =="South Carolina"| State =="Tennessee"| State =="Texas"| State =="Virgina"
  	   | State =="West Virgina")
pacific<- project_data %>%
  filter(State == "Alaska"| State == "Hawaii")
```

**Regions are composed as West:**
```{r, echo=FALSE}
str(west$State)
```
**Midwest**
```{r, echo=FALSE}
str(midwest$State)
```
**Northeast**
```{r, echo=FALSE}
str(northeast$State)
```
**South**
```{r, echo=FALSE}
str(south$State)
```
**&  Pacific**
```{r, echo=FALSE}
str(pacific$State)
```


![US regions based on [census.gov](https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf) ](regions.PNG){width=50%}


### Research question 

**Question: What region in the United States have the worst drivers?**

$H_0$ = $Region1_{fatal.accidents}\le Region(2...5)_{fatal.accidents}$
$H_A$ = $Region1_{fatal.accidents}> Region(2...5)_{fatal.accidents}$




### Cases 

**What are the cases, and how many are there?**



### Data collection 

**Describe the method of data collection.**



### Type of study 

**What type of study is this (observational/experiment)?**



### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**




### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**



### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**



### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}

```

